---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# fread-benchmarks-rsuite

<!-- badges: start -->
<!-- badges: end -->

## Introduction
This is the `rsuite` version of teh original tests published by [Jozef Hajnala](https://gitlab.com/jozefhajnala) in [Gitlab](https://gitlab.com/jozefhajnala/fread-benchmarks/tree/master). 

It is conceptually the same scripts but adding these changes:

1. a new R script using `awk`
2. modification of the project structure using the `rsuite` paradigm
3. addition of comments on scripts
4. generation of results dataframe
5. plots for memory and average time per operation
6. packages are all installed exclusively for this project, under the folder `packages`, insulated from the global environment

## Data
The data is the same used in the original project (airlines), CSV files located [here](http://stat-computing.org/dataexpo/2009/the-data.html). The files are downloaded compressed as bunzip, then expanded as `csv`.


## Running the benchmarks

### For base R

```
bash bench.sh R/01_base.R &> results/out_base.txt
```

### For data.table::fread

```
bash bench.sh R/02_fread.R &> results/out_fread.txt
```

### For readr::read_csv

```
bash bench.sh R/03_readr.R &> results/out_readr.txt
```

### For data.table::fread with grep

```
bash bench.sh R/04_fread_grep.R &> results/out_fread_grep.txt
```


### For readr::read_csv with grep

```
bash bench.sh R/05_readr_grep.R &> results/out_readr_grep.txt
```

### For data.table::fread with awk

```
bash bench.sh R/04_fread_grep.R &> results/out_fread_awk.txt
```


## Results

### Tables

```{r message=FALSE}
library(dplyr)

source("R/show_results.R")

df <- get_results_df() %>% 
    arrange(-mem_gb) %>% 
    select(result_file, rscript, mem_gb, avg_secs) %>% 
    mutate_if(is.numeric, round, digits = 2) %>% 
    print()
```

This is the same table but including the original description for the scripts.

```{r echo=FALSE}
source("R/show_results.R")
get_results_summary(col_names = c("description", "rscript", "mem_gb", "avg_secs")) %>% 
    mutate_if(is.numeric, round, digits = 2)
```

- mem_gb = Maximum resident set size, gigabytes
- avg_secs = Average of real time and user time as measured by `time`, seconds

### Memory required (GB)

```{r}
library(ggplot2)
library(scales)

df$rscript <- reorder(df$rscript, df$mem_gb)
ggplot(df, aes(x = rscript, y = mem_gb)) +
    geom_col() +
    scale_y_continuous(breaks=pretty_breaks(n=20)) +
    coord_flip() 
```

### Average Time (seconds)

```{r}
df$rscript <- reorder(df$rscript, df$avg_secs)
ggplot(df, aes(x = rscript, y = avg_secs)) +
    geom_col() +
    scale_y_continuous(breaks=pretty_breaks(n=20)) +
    coord_flip()
```

## How to reproduce this project yourself
1. Download and install the [RSuite client](https://rsuite.io/RSuite_Download.php). Available for Linux, Mac and Windows.
2. Install the `rsuite` package with `rsuite install`
3. Clone or download this repository.
4. Change to this repo folder and install the dependencies on its own isolated reproducible environment. Use `rsuite proj depsinst`
5. Build the project with `rsuite proj build`
6. Download the data running this from the console `Rscript R/data_prep.R`
7. Run each of the tests. Example: `bash bench.sh R/01_base.R &> results/out_base.txt`. See above for the rest.
8. Enjoy


## References

* Original article by [Jozef Hajnala](https://twitter.com/jozefhajnala): [How data.table's fread can save you a lot of time and memory, and take input from shell commands](https://jozef.io/r917-fread-comparisons/)

* Original repository: https://gitlab.com/jozefhajnala/fread-benchmarks

* Article by [Nick Strayer](https://twitter.com/NicholasStrayer): [Using AWK and R to parse 25tb](https://livefreeordichotomize.com/2019/06/04/using_awk_and_r_to_parse_25tb/)

* This [repository]() by [Alfonso R. Reyes](https://twitter.com/OilGains)

* [Rsuite client](https://rsuite.io/)
